---
title: "Aortic valve analysis"
author: "Steve Simon"
date: "May 25, 2017"
output: html_document
---


```{r preliminaries, echo=FALSE}
library(dplyr)
library(lubridate)
library(magrittr)
library(survival)
```

```{r read-data}
av1 <- read.csv("aorta1.csv", as.is=TRUE)
av3 <- read.csv("aorta3.csv", as.is=TRUE)
av4 <- read.csv("aorta4.csv", as.is=TRUE)

print(names(av1)[1])
names(av1)[1] <- "id"
print(names(av3)[1])
names(av3)[1] <- "id"
print(names(av4)[1])
names(av4)[1] <- "id"

names(av1) %<>% tolower
names(av3) %<>% tolower
names(av4) %<>% tolower

names(av1) %<>% gsub("\\.", "_", .)
names(av3) %<>% gsub("\\.", "_", .)
names(av4) %<>% gsub("\\.", "_", .)

names(av1) %<>% gsub("_+",  "_", .)
names(av3) %<>% gsub("_+",  "_", .)
names(av4) %<>% gsub("_+",  "_", .)

names(av1) %<>% gsub("_$",  "",  .)
names(av3) %<>% gsub("_$",  "",  .)
names(av4) %<>% gsub("_$",  "",  .)
```

AORTIC
 
1.	Arithmetic data for columns where possible – mean, median, SD, range

2.	What is the long-term survival of children with AVR?

3.	Head to head comparison of Ross versus Bio prosthesis versus Mechanical prosthesis looking at survival within the 3 groups and relation to following
 
* Age at 1st AVR
* Weight at 1st AVR
* Valve type at 1st AVR
* Valve size at 1st AVR
* Underlying disease
* Underlying hemodynamic mechanism – AR versus AS versus both
* Perfusion time
* Concomitant prosthetic MVR
* Initial Hospital length of stay
* Era effect - < 1990 versus 1990s and up to 2003
* Geography – State 
* Institution

4.	Causes of death – Big piles only

5.	Early deaths (PCCC) versus late deaths (NDI) – compare profiles

6.	Interactions

7.	Confounders – collinearity

8.	Kaplan Meir – Ross versus Bio versus Mechanical AVR

9.	Competing risk – will think thru this some more

10.	Redo stuff – we may not have time for this before abstract deadline


```{r peek}
j1 <- which(grepl("^x",names(av1)))
print(names(av1)[j1])
j1 <- j1[-1]
j3 <- which(grepl("^x",names(av3)))
print(j3)
j4 <- which(grepl("^x",names(av4)))
print(names(av4)[j4])

for (j in names(av1)[j1]) {
  x <- av1[, j]
  n <- length(x)
  is_empty <- is.na(x) | grepl("^[[:space:]]+$", x)
  if (sum(is_empty)==n) {next}
  cat("\n\n")
  cat(j)
  cat("\n")
  print(table(x[!is_empty]))
}

i1 <- which(is.na(av1$id))
print(i1)
i3 <- which(is.na(av3$id))
print(i3)
i4 <- which(is.na(av4$id))
print(i4)

for (j in names(av3)) {
  x <- av3[i3, j]
  n <- length(x)
  is_empty <- is.na(x) | x=="" | grepl("^[[:space:]]+$", x)
  if (sum(is_empty)==n) {next}
  cat("\n\n")
  cat(j)
  cat(": ")
  cat(i3[!is_empty])
  print(table(x[!is_empty]))
}

av1 <- av1[, -j1]

av3 <- av3[-i3, ]

str(av1)
head(av1)
tail(av1)

str(av3)
head(av3)
tail(av3)

str(av4)
head(av4)
tail(av4)
```

```{r examine-dates}
isolate_month <- function(v) {
  v                                             %>%
    gsub("/.*", "", .)                          %>%
    gsub("-.*", "", .)                          %>%
    return
}
isolate_day <- function(v) {
  v                                             %>%
    sub("^.*?/", "", .)                         %>%
    sub("/.*$?", "", .)                         %>%
    sub("^.*?-", "", .)                         %>%
    sub("-.*$?", "", .)                         %>%
    return
}
isolate_year <- function(v) {
  v                                             %>%
    sub("^.*/", "", .)                          %>%
    sub("^.*-", "", .)                          %>%
    return
}
examine_dates <- function(v) {
  two_slashes <- grepl("/", sub("/", "", v))
  two_dashes  <- grepl("-", sub("-", "", v))
  special_cases <- 
    ! (two_slashes | two_dashes) |
    is.na(v)
  if (sum(special_cases)>0) {
    cat("\n\n")
    cat("Special cases")
    v[special_cases]                            %>%
      table(useNA="ifany")                      %>%
      addmargins                                %>%
      print
  }
  cat("\n\n")
  cat("Month")
  v[!special_cases]                             %>%
    isolate_month                               %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
  cat("\n\n")
  cat("Day")
  v[!special_cases]                             %>%
    isolate_day                                 %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
  cat("\n\n")
  cat("Year")
  v[!special_cases]                             %>%
    isolate_year                                %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
}
alist1 <- c(
  "birthdate",
  "first_admit_date",
  "date_of_valve_replacement",
  "discharge_date",
  "date_of_valve_replacement_1",
  "discharge_date_1",
  "date_of_valve_replacement_2",
  "discharge_date_2",
  "unos_init_date",
  "unos_end_date",
  "unos_tx_date",
  "unos_re_tx_date",
  "pccc_tx_date",
  "ndi_date_of_death")
for (a in alist1) {
  cat("\n\n")
  cat(a)
  examine_dates(av1[, a]) 
}
{
  cat("\n\n")
  cat("date_of_death")
  examine_dates(av3$date_of_death)
}
{
  cat("\n\n")
  cat("date_of_death")
  examine_dates(av4$date_of_death)
}
```

```{r merge-everything}
print(length(unique(av1$id)))
print(dim(av1)[1])
print(length(unique(av3$id)))
print(dim(av3)[1])
print(length(unique(av4$id)))
print(dim(av4)[1])

print(length(intersect(av1$id, av3$id)))
print(length(intersect(av1$id, av4$id)))
print(length(intersect(av3$id, av4$id)))
print(setdiff(av3$id, av1$id))

av1                                             %>%
  select(id, ndi_date_of_death)                 %>%
  inner_join(av3)                               %>%
  mutate(n=(ndi_date_of_death==date))
```

```{r save-everything, echo=FALSE}
save.image("av.RData")
```