---
title: "Aortic valve analysis"
author: "Steve Simon"
date: "May 25, 2017"
output: html_document
---


```{r preliminaries, echo=FALSE}
source("prelims.R", echo=FALSE)
av1 <- read.csv("av1_2017_05_29.csv", as.is=TRUE)
av3 <- read.csv("av3_2017_05_29.csv", as.is=TRUE)
av4 <- read.csv("av4_2017_05_29.csv", as.is=TRUE)

mv1 <- read.csv("mv1_2017_05_29.csv", as.is=TRUE)
mv4 <- read.csv("mv4_2017_05_29.csv", as.is=TRUE)


names(av1) %<>% clean_up_names
names(av3) %<>% clean_up_names
names(av4) %<>% clean_up_names

names(mv1) %<>% clean_up_names
names(mv4) %<>% clean_up_names

av1 <- remove_empty_space(av1)
av3 <- remove_empty_space(av3)
av4 <- remove_empty_space(av4)

mv1 <- remove_empty_space(mv1)
mv4 <- remove_empty_space(mv4)

loop_through_dates(av1)
loop_through_dates(av3)
loop_through_dates(av4)

loop_through_dates(mv1)
loop_through_dates(mv4)
```

AORTIC
 
1.	Arithmetic data for columns where possible – mean, median, SD, range

2.	What is the long-term survival of children with AVR?

3.	Head to head comparison of Ross versus Bio prosthesis versus Mechanical prosthesis looking at survival within the 3 groups  ==>
`r names(av1)[23]` and relation to following
 
* Age at 1st AVR ==> 
`r names(av1)[16]` MINUS
`r names(av1)[ 2]`
* Weight at 1st AVR ==>
`r names(av1)[18]`
* Valve type at 1st AVR ==>
`r names(av1)[14]` AND 
`r names(av1)[20]`
* Valve size at 1st AVR ==>
`r names(av1)[19]`
* Underlying disease ==> 
`r names(av1)[13]`
* Underlying hemodynamic mechanism – AR versus AS versus both ==> 
`r names(av1)[15]`
* Perfusion time ==> 
`r names(av1)[21]`
* Concomitant prosthetic MVR ==> 
`r names(av1)[22]`
* Initial Hospital length of stay ==> 
`r names(av1)[17]` MINUS 
`r names(av1)[16]`
* Era effect - < 1990 versus 1990s and up to 2003 ==>
`r names(av1)[16]`
* Geography – State ==> 
`r names(av1)[6]`
* Institution

4.	Causes of death – Big piles only

5.	Early deaths (PCCC) versus late deaths (NDI) – compare profiles

6.	Interactions

7.	Confounders – collinearity

8.	Kaplan Meir – Ross versus Bio versus Mechanical AVR

9.	Competing risk – will think thru this some more

10.	Redo stuff – we may not have time for this before abstract deadline


```{r peek}
j1 <- which(grepl("^x",names(av1)))
print(names(av1)[j1])
j1 <- j1[-1]
j3 <- which(grepl("^x",names(av3)))
print(j3)
j4 <- which(grepl("^x",names(av4)))
print(names(av4)[j4])

for (j in names(av1)[j1]) {
  x <- av1[, j]
  n <- length(x)
  is_empty <- is.na(x) | grepl("^[[:space:]]+$", x)
  if (sum(is_empty)==n) {next}
  cat("\n\n")
  cat(j)
  cat("\n")
  print(table(x[!is_empty]))
}

i1 <- which(is.na(av1$id))
print(i1)
i3 <- which(is.na(av3$id))
print(i3)
i4 <- which(is.na(av4$id))
print(i4)

for (j in names(av3)) {
  x <- av3[i3, j]
  n <- length(x)
  is_empty <- is.na(x) | x=="" | grepl("^[[:space:]]+$", x)
  if (sum(is_empty)==n) {next}
  cat("\n\n")
  cat(j)
  cat(": ")
  cat(i3[!is_empty])
  print(table(x[!is_empty]))
}

av1 <- av1[, -j1]

av3 <- av3[-i3, ]

str(av1)
head(av1)
tail(av1)

str(av3)
head(av3)
tail(av3)

str(av4)
head(av4)
tail(av4)
```

```{r examine-dates}
alist1 <- c(
  "birthdate",
  "first_admit_date",
  "date_of_valve_replacement",
  "discharge_date",
  "date_of_valve_replacement_1",
  "discharge_date_1",
  "date_of_valve_replacement_2",
  "discharge_date_2",
  "unos_init_date",
  "unos_end_date",
  "unos_tx_date",
  "unos_re_tx_date",
  "pccc_tx_date",
  "ndi_date_of_death")
for (a in alist1) {
  cat("\n\n")
  cat(a)
  examine_dates(av1[, a]) 
}
{
  cat("\n\n")
  cat("date_of_death")
  examine_dates(av3$date_of_death)
}
{
  cat("\n\n")
  cat("date_of_death")
  examine_dates(av4$date_of_death)
}
```

```{r merge-everything}
print(length(unique(av1$id)))
print(dim(av1)[1])
print(length(unique(av3$id)))
print(dim(av3)[1])
print(length(unique(av4$id)))
print(dim(av4)[1])

print(length(intersect(av1$id, av3$id)))
print(length(intersect(av1$id, av4$id)))
print(length(intersect(av3$id, av4$id)))
print(setdiff(av3$id, av1$id))

av1                                             %>%
  select(id, ndi_date_of_death)                 %>%
  inner_join(av3)                               %>%
  mutate(n=(ndi_date_of_death!=date_of_death))  %>%
  use_series(n)                                 %>%
  sum
av1                                             %>%
  select(id, ndi_date_of_death)                 %>%
  anti_join(av3)                                %>%
  use_series(ndi_date_of_death)                 %>%
  table
```

```{r fix_dates}
fix_list1 <- "date_of_valve_replacement"
for (d in fix_list1) {
  v <- av1[, d]
  double_delimiter <-  grepl("[-/].*[-/]", v)
  single_delimiter <-  grepl("[-/]", v) & ! double_delimiter
  absent_delimiter <- !grepl("[-/]", v)
  year_only <- absent_delimiter & as.numeric(v) %in% as.character(1950:2020)
  missing_code <- absent_delimiter & !year_only
  double_00 <- grepl("00/00", v) | grepl("^0/00", v)
  v <- sub("^0/00", "00/00", v)
  single_00 <- grepl("/00/",  v) & !double_00 
  cat("\n\nMissing code\n")
  v[missing_code] %>% unique %>% sort %>% print
  cat("\n\nYear only\n")
  v[year_only]        %>% unique %>% sort %>% print
  cat("\n\nSingle delimiter\n")
  v[single_delimiter] %>% unique %>% sort %>% print
  cat("\n\nDouble 00\n")
  v[double_00]        %>% unique %>% sort %>% print
  cat("\n\nSingle 00\n")
  v[single_00]        %>% unique %>% sort %>% print
  cat("\n\n")

  w <- v
  w[year_only] <- paste0("06/30/", v[year_only])
  w[single_delimiter] <- sub("/", "/15/", v[single_delimiter])
  w[single_00] <- sub("/00/", "/15/", v[single_00])
  w[double_00] <- sub("^0/00", "06/30", v[double_00])
  w[double_00] <- sub("00/00", "06/30", v[double_00])
  print(data.frame(changes=paste(v, "changes to", w)[v!=w]))
  av1$date_of_valve_replacement_recode <- w
}
```

```{r kaplan-meier}
# Check for dates that will not convert.
m_codes <- av1$date_of_valve_replacement_recode %in% c("888", "999", "999.00")
sum(m_codes)
av1 %>% filter(!m_codes) -> av1x
dim(av1x)
d0 <- mdy(av1x$date_of_valve_replacement_recode) - mdy("1/1/1950")
d1 <- mdy(av1x$ndi_date_of_death)                - mdy("1/1/1950")
d2 <- rep(mdy("12/30/2014")-mdy("1/1/1950"), dim(av1x)[1])
sum(is.na(d0))
sum(is.na(d1))
c1 <- as.numeric(d1-d0) / 365.25
c2 <- as.numeric(d2-d0) / 365.25
censor <- as.numeric(is.finite(c1))
table(censor)
c1[censor==0] <- c2[censor==0]
av_su <- Surv(time=round(c1, 4), event=censor)
plot(survfit(av_su~1))
```


```{r save-everything, echo=FALSE}
save.image("av.RData")
```