---
title: "Aortic valve analysis"
author: "Steve Simon"
date: "May 25, 2017"
output: html_document
---


```{r preliminaries}
source("prelims.R", echo=FALSE)
av1 <- read.csv("av1_2017_05_29.csv", as.is=TRUE)
av3 <- read.csv("av3_2017_05_29.csv", as.is=TRUE)
av4 <- read.csv("av4_2017_05_29.csv", as.is=TRUE)

mv1 <- read.csv("mv1_2017_05_29.csv", as.is=TRUE)
mv4 <- read.csv("mv4_2017_05_29.csv", as.is=TRUE)


names(av1) %<>% clean_up_names
names(av3) %<>% clean_up_names
names(av4) %<>% clean_up_names

names(mv1) %<>% clean_up_names
names(mv4) %<>% clean_up_names

av1 <- remove_empty_space(av1)
av3 <- remove_empty_space(av3)
av4 <- remove_empty_space(av4)

mv1 <- remove_empty_space(mv1)
mv4 <- remove_empty_space(mv4)

av1 <- check_ids(av1)
av3 <- check_ids(av3)
av4 <- check_ids(av4)

mv1 <- check_ids(mv1)
mv4 <- check_ids(mv4)

{
  cat("\n\n\n\n**** Note new codes: ")
  cat("998 (was 888), ")
  cat("997 (was blank), ")
  cat("996 (was missing), ")
  cat("995 (unfixable)")
  cat("\n\n")
}

av1 <- loop_through_dates(av1)
av3 <- loop_through_dates(av3)
av4 <- loop_through_dates(av4)

mv1 <- loop_through_dates(mv1)
mv4 <- loop_through_dates(mv4)

print(dim(av1)[1])
print(dim(av3)[1])
print(dim(av4)[1])

print(length(intersect(av1$id, av3$id)))
print(length(intersect(av1$id, av4$id)))
print(length(intersect(av3$id, av4$id)))
print(setdiff(av3$id, av1$id))

print(dim(mv1)[1])
print(dim(mv4)[1])

print(length(intersect(mv1$id, mv4$id)))
print(setdiff(mv4$id, mv1$id))
```

AORTIC
 
1.	Arithmetic data for columns where possible – mean, median, SD, range

2.	What is the long-term survival of children with AVR?

3.	Head to head comparison of Ross versus Bio prosthesis versus Mechanical prosthesis looking at survival within the 3 groups  ==>
`r names(av1)[23]` and relation to following
 
* Age at 1st AVR ==> 
`r names(av1)[16]` MINUS
`r names(av1)[ 2]`
* Weight at 1st AVR ==>
`r names(av1)[18]`
* Valve type at 1st AVR ==>
`r names(av1)[14]` AND 
`r names(av1)[20]`
* Valve size at 1st AVR ==>
`r names(av1)[19]`
* Underlying disease ==> 
`r names(av1)[13]`
* Underlying hemodynamic mechanism – AR versus AS versus both ==> 
`r names(av1)[15]`
* Perfusion time ==> 
`r names(av1)[21]`
* Concomitant prosthetic MVR ==> 
`r names(av1)[22]`
* Initial Hospital length of stay ==> 
`r names(av1)[17]` MINUS 
`r names(av1)[16]`
* Era effect - < 1990 versus 1990s and up to 2003 ==>
`r names(av1)[16]`
* Geography – State ==> 
`r names(av1)[6]`
* Institution

4.	Causes of death – Big piles only
5.	Early deaths (PCCC) versus late deaths (NDI) – compare profiles

6.	Interactions

7.	Confounders – collinearity

8.	Kaplan Meir – Ross versus Bio versus Mechanical AVR

9.	Competing risk – will think thru this some more

10.	Redo stuff – we may not have time for this before abstract deadline

```{r calculate-dates}
m_codes <- as.character(995:999)

calculate_days <- function(df, v) {
  dfname <- deparse(substitute(df))
  vname <- deparse(substitute(v))
  t0 <- mdy("1/1/1950") # our overall reference point
  missing_codes <- df[, vname] %in% m_codes
  cat("\nThere are ")
  cat(dim(df)[1]-sum(missing_codes))
  cat(" valid values out of ")
  cat(dim(df)[1])
  cat(" for ")
  cat(paste(dfname, vname, sep="$"))
  cat(".\n")
  days <- mdy(df[, vname])-t0
  if (sum(is.na(days)>sum(missing_codes))) {
    cat("\nValues for which conversion was not possible\n")
    df[is.na(days)&!missing_codes, vname] %>% table %>% print
  }
  return(days)
}

av3                                             %>%
  bind_rows(av4)                                %>%
  filter(!(date_of_death %in% m_codes))         %>%
  select(id, date_of_death)                     %>%
  right_join(av1)                               -> av5  
av5[is.na(av5$date_of_death), "date_of_death"] <- 996

av5$d_birth <- calculate_days(av5, birthdate)
av5$d_vr1   <- calculate_days(av5, date_of_valve_replacement)
av5$d_vr2   <- calculate_days(av5, date_of_valve_replacement_1)
av5$d_vr3   <- calculate_days(av5, date_of_valve_replacement_2)
av5$d_dis1  <- calculate_days(av5, discharge_date)
av5$d_dis2  <- calculate_days(av5, discharge_date_1)
av5$d_dis3  <- calculate_days(av5, discharge_date_2)
av5$d_tx    <- calculate_days(av5, unos_tx_date)
av5$d_death <- calculate_days(av5, date_of_death)
av5$d_cens  <- mdy("12/30/2014") - mdy("1/1/1950")

mv4                                             %>%
  filter(!(date_of_death %in% m_codes))         %>%
  select(id, date_of_death)                     %>%
  right_join(mv1)                               -> mv5  
mv5[is.na(mv5$date_of_death), "date_of_death"] <- 996

mv5$d_birth <- calculate_days(mv5, birthdate)
mv5$d_vr1   <- calculate_days(mv5, date_of_valve_replacement)
mv5$d_vr2   <- calculate_days(mv5, date_of_valve_replacement_1)
mv5$d_vr3   <- calculate_days(mv5, date_of_valve_replacement_2)
mv5$d_dis1  <- calculate_days(mv5, discharge_date)
mv5$d_dis2  <- calculate_days(mv5, discharge_date_1)
mv5$d_dis3  <- calculate_days(mv5, discharge_date_2)
mv5$d_tx    <- calculate_days(mv5, unos_tx_date)
mv5$d_death <- calculate_days(mv5, date_of_death)
mv5$d_cens  <- mdy("12/30/2014") - mdy("1/1/1950")
```

```{r mortality-only}
t1 <- av5$d_death-av5$d_birth
t2 <- av5$d_cens -av5$d_birth
t <- ifelse(is.finite(t1), t1, t2)
cens <- as.numeric(is.finite(t1))
av_mortality <- Surv(time=round(t/365.25, 4), event=cens)
plot(survfit(av_mortality~1))

t1 <- mv5$d_death-mv5$d_birth
t2 <- mv5$d_cens -mv5$d_birth
t <- ifelse(is.finite(t1), t1, t2)
cens <- as.numeric(is.finite(t1))
mv_mortality <- Surv(time=round(t/365.25, 4), event=cens)
plot(survfit(mv_mortality~1))
```

```{r conpeting-risk-second-replacement}
# Several cases to consider

# Case #1, alive at the end of the study, no second replacement
# Case #2, died without second replacement
# Case #3, alive at the end of the study, second replacement procedure needed
# Case #4, died after second replacement procedure

table(is.na(d_death), is.na(d_vr2))

```

```{r save-everything, echo=FALSE}
save.image("clean_up.RData")
```