---
title: "Mitral valve analysis"
author: "Steve Simon"
date: "May 23, 2017"
output: html_document
---


```{r preliminaries, echo=FALSE}
library(dplyr)
library(lubridate)
library(magrittr)
library(survival)
```

```{r read-data}
mv1 <- read.csv("mitral1.csv", as.is=TRUE)
mv3 <- read.csv("mitral3.csv", as.is=TRUE)
names(mv1) %<>% tolower
names(mv3) %<>% tolower
names(mv1) %<>% gsub("\\.", "_", .)
names(mv3) %<>% gsub("\\.", "_", .)
names(mv1) %<>% gsub("_+",  "_", .)
names(mv3) %<>% gsub("_+",  "_", .)
names(mv1) %<>% gsub("_$",  "",  .)
names(mv3) %<>% gsub("_$",  "",  .)
```


MITRAL
 
1. Arithmetic data for columns where possible – mean, median, SD, range

2. What is the long-term survival in children who have had prosthetic MVR?

3. Relationship of survival to
* Age at 1st MVR (P-B)
`r names(mv1)[2]`
`r names(mv1)[16]`
* Weight at 1st MVR (Q)
`r names(mv1)[17]`
* Valve type at 1st MVR (N, but also S)
`r names(mv1)[14]`
`r names(mv1)[19]`
* Valve size at 1st MVR (R)
`r names(mv1)[18]`
* Underlying disease (M)
`r names(mv1)[13]`
* Underlying hemodynamic mechanism – MR versus MS versus both (O)
`r names(mv1)[15]`
* Perfusion time (T)
`r names(mv1)[20]`
* Concomitant prosthetic AVR (V)
`r names(mv1)[22]`
* Initial Hospital length of stay (U-P)
`r names(mv1)[21]` MINUS 
`r names(mv1)[16]`
* Era effect - < 1990 versus 1990s and up to 2003 (P)
`r names(mv1)[16]`
* Geography – State (F)
`r names(mv1)[6]`
* Institution (not important)

4.	Causes of death – big piles only (Tab 3, D and E)
`r names(mv3)[4]` AND
`r names(mv3)[5]`

5.	Early deaths (PCCC) versus late deaths (NDI) – compare profiles (H versus BI)
`r names(mv1)[8]` AND
`r names(mv1)[61]`

6.	Interactions (think about it)

7.	Confounders – collinearity

8.	Need competing risk plot – I need to think about what goes in this once we see some results

9.	Redo stuff – we may not have time for this before abstract deadline
 

```{r peek}
i1 <- which(names(mv1)=="x")
i2 <- dim(mv1)[2]
for (i in i1:i2) {
  print(table(mv1[, i], useNA="always"))
}
j <- which(is.na(mv1$patient_id))
mv1[j, ]
mv1 <- mv1[-j, 1:(i1-1)]

str(mv1)
head(mv1)
tail(mv1)

str(mv3)
head(mv3)
tail(mv3)
```

```{r examine-dates}
isolate_month <- function(v) {
  v                                             %>%
    gsub("/.*", "", .)                          %>%
    gsub("-.*", "", .)                          %>%
    return
}
isolate_day <- function(v) {
  v                                             %>%
    sub("^.*?/", "", .)                         %>%
    sub("/.*$?", "", .)                         %>%
    sub("^.*?-", "", .)                         %>%
    sub("-.*$?", "", .)                         %>%
    return
}
isolate_year <- function(v) {
  v                                             %>%
    sub("^.*/", "", .)                          %>%
    sub("^.*-", "", .)                          %>%
    return
}
examine_dates <- function(v) {
  two_slashes <- grepl("/", sub("/", "", v))
  two_dashes  <- grepl("-", sub("-", "", v))
  special_cases <- 
    ! (two_slashes | two_dashes) |
    is.na(v)
  if (sum(special_cases)>0) {
    cat("\n\n")
    cat("Special cases")
    v[special_cases]                            %>%
      table(useNA="ifany")                      %>%
      addmargins                                %>%
      print
  }
  cat("\n\n")
  cat("Month")
  v[!special_cases]                             %>%
    isolate_month                               %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
  cat("\n\n")
  cat("Day")
  v[!special_cases]                             %>%
    isolate_day                                 %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
  cat("\n\n")
  cat("Year")
  v[!special_cases]                             %>%
    isolate_year                                %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
}
dlist1 <- c(
  "birthdate",
  "first_admit_date",
  "date_of_valve_replacement",
  "discharge_date",
  "date_of_valve_replacement_1",
  "discharge_date_1",
  "date_of_valve_replacement_2",
  "discharge_date_2",
  "unos_init_date",
  "unos_end_date",
  "unos_tx_date",
  "unos_re_tx_date",
  "pccc_tx_date",
  "ndi_date_of_death")
for (d in dlist1) {
  cat("\n\n")
  cat(d)
  examine_dates(mv1[, d]) 
}
{
  cat("\n\n")
  cat("date_of_death")
  examine_dates(mv3$date_of_death)
}
```

```{r fix_dates}
fix_list1 <- "date_of_valve_replacement"
for (d in fix_list1) {
  v <- mv1[, d]
  double_delimiter <-  grepl("[-/].*[-/]", v)
  single_delimiter <-  grepl("[-/]", v) & ! double_delimiter
  absent_delimiter <- !grepl("[-/]", v)
  year_only <- absent_delimiter & as.numeric(v) %in% as.character(1950:2020)
  missing_code <- absent_delimiter & !year_only
  double_00 <- grepl("00/00", v)
  single_00 <- grepl("/00/",  v) & !double_00 
  cat("\n\nMissing code\n")
  v[missing_code] %>% unique %>% sort %>% print
  cat("\n\nYear only\n")
  v[year_only]        %>% unique %>% sort %>% print
  cat("\n\nSingle delimiter\n")
  v[single_delimiter] %>% unique %>% sort %>% print
  cat("\n\nDouble 00\n")
  v[double_00]        %>% unique %>% sort %>% print
  cat("\n\nSingle 00\n")
  v[single_00]        %>% unique %>% sort %>% print
  cat("\n\n")

  w <- v
  w[year_only] <- paste0("06/30/", v[year_only])
  w[single_delimiter] <- sub("/", "/15/", v[single_delimiter])
  w[single_00] <- sub("/00/", "/15/", v[single_00])
  w[double_00] <- sub("00/00", "06/30", v[double_00])
  print(data.frame(changes=paste(v, "changes to", w)[v!=w]))
  mv1$date_of_valve_replacement_recode <- w
}
```

```{r merge}
dim(mv1)
length(unique(mv1$patient_id))
dim(mv3)
length(unique(mv3$patientid))
length(intersect(mv1$patient_id, mv3$patientid))
length(setdiff(mv1$patient_id, mv3$patientid))
length(setdiff(mv3$patientid, mv1$patient_id))
setdiff(mv3$patientid, mv1$patient_id)
mv3                                             %>%
  rename(patient_id=patientid)                  %>%
  select(patient_id, date_of_death)            %>%
  inner_join(mv1)                               %>%
  select(patient_id, ndi_date_of_death, date_of_death) -> mv4
dim(mv4)
sum(mv4$ndi_date_of_death != mv4$date_of_death)
sum(mv1$ndi_date_of_death == "")
mv3                                             %>%
  rename(patient_id=patientid)                  %>%
  select(patient_id, date_of_death)            %>%
  inner_join(mv1)                               %>%
  select(patient_id, ndi_date_of_death, date_of_death) -> mv4
dim(mv4)
sum(mv4$ndi_date_of_death != mv4$date_of_death)
sum(mv1$ndi_date_of_death == "")
```

```{r descriptives}
v_list1 <- c(
  "weight_at_surgery_kg",
  "cpb_time_minutes"
)
for (i in v_list1) {
  n888 <- sum(mv1[, i]==888)
  n999 <- sum(mv1[, i]==999)
  mv1[, i] <- ifelse(mv1[, i] %in% c(888, 999), NA, mv1[, i])
  cat("\n\n")
  cat(i)
  cat("\n")
  cat(paste("\nCount of 888 =", n888))
  cat(paste("\nCount of 999 =", n999))
  cat("\n")
  print(summary(mv1[, i]))
}
v_list2 <- c(
  "mechanical_valve_used_y_n",
  "valve_make",
  "valve_size_mm",
  "for_all_columns_999_unknown_888_n_a_underlying_disease_ex_cavc_etc",
  "underlying_lesion_regurgitation_or_stenosis",
  "additional_valve_replaced_aortic_pulm_tri",
  "state"
)
for (i in v_list2) {
  n888 <- sum(mv1[, i]=="888")
  n999 <- sum(mv1[, i]=="999")
  mv1[, i] %<>% sub("^ +", "", .)
  mv1[, i] %<>% sub(" +$", "", .)
  mv1[, i] %<>% sub(rawToChar(as.raw(160)), "", .)
  mv1[, i] <- ifelse(mv1[, i] %in% c("888", "999"), NA, mv1[, i])
  cat("\n\n")
  cat(i)
  cat("\n")
  cat(paste("\nCount of 888 =", n888))
  cat(paste("\nCount of 999 =", n999))
  cat("\n")
  print(table(mv1[, i], useNA="always"))
}
```

```{r kaplan-meier}
# Check for dates that will not convert.
m_codes <- mv1$date_of_valve_replacement_recode %in% c("888", "999")
sum(m_codes)
mv1 %>% filter(!m_codes) -> mv1x
dim(mv1x)
d0 <- mdy(mv1x$date_of_valve_replacement_recode) - mdy("1/1/1950")
d1 <- mdy(mv1x$ndi_date_of_death)                - mdy("1/1/1950")
d2 <- rep(mdy("12/30/2014")-mdy("1/1/1950"), dim(mv1x)[1])
sum(is.na(d0))
sum(is.na(d1))
c1 <- as.numeric(d1-d0) / 365.25
c2 <- as.numeric(d2-d0) / 365.25
censor <- as.numeric(is.finite(c1))
table(censor)
c1[censor==0] <- c2[censor==0]
mv_su <- Surv(time=round(c1, 4), event=censor)
plot(survfit(mv_su~1))
```

```{r save-everything, echo=FALSE}
save.image("mv.RData")
```