---
title: "Mitral valve analysis"
author: "Steve Simon"
date: "May 23, 2017"
output: html_document
---

```{r preliminaries, echo=FALSE}
library(dplyr)
library(lubridate)
library(magrittr)
```

```{r read-data}
mv1 <- read.csv("mitral1.csv", as.is=TRUE)
mv3 <- read.csv("mitral3.csv", as.is=TRUE)
names(mv1) %<>% tolower
names(mv3) %<>% tolower
names(mv1) %<>% gsub("\\.",   "_", .)
names(mv3) %<>% gsub("\\.",   "_", .)
names(mv1) %<>% gsub("_+",    "_", .)
names(mv3) %<>% gsub("_+",    "_", .)
i1 <- which(names(mv1)=="x")
i2 <- dim(mv1)[2]
for (i in i1:i2) {
  print(table(mv1[, i], useNA="always"))
}
j <- which(is.na(mv1$patient_id))
mv1[j, ]
mv1 <- mv1[-j, 1:(i1-1)]
str(mv1)
head(mv1)
tail(mv1)
str(mv3)
head(mv3)
tail(mv3)
```


MITRAL
 
1.	Arithmetic data for columns where possible – mean, median, SD, range
2.	What is the long-term survival in children who have had prosthetic MVR?
3.	Relationship of survival to
a.  Age at 1st MVR (P-B) `r names(mv1)[2]` and `r names(mv1)[16]`
b.  Weight at 1st MVR (Q) `r names(mv1)[17]`
c.  Valve type at 1st MVR (N, but also S) `r names(mv1[c(14, 19)])`
d.  Valve size at 1st MVR (R)
e.  Underlying disease (M)
f.  Underlying hemodynamic mechanism – MR versus MS versus both (O)
g.  Perfusion time (T)
h.  Concomitant prosthetic AVR (V)
i.  Initial Hospital length of stay (U-P)
j.  Era effect - < 1990 versus 1990s and up to 2003 (P)
k.  Geography – State (F)
l.  Institution (not important)
4.	Causes of death – big piles only (Tab 3, D and E)
5.	Early deaths (PCCC) versus late deaths (NDI) – compare profiles (H versus BI)
6.	Interactions (think about it)
7.	Confounders – collinearity
8.	Need competing risk plot – I need to think about what goes in this once we see some results
9.	Redo stuff – we may not have time for this before abstract deadline
 
AORTIC
 
1.	Arithmetic data for columns where possible – mean, median, SD, range
2.	What is the long-term survival of children with AVR?
3.	Head to head comparison of Ross versus Bio prosthesis versus Mechanical prosthesis looking at survival within the 3 groups and relation to following
 
Age at 1st AVR
Weight at 1st AVR
Valve type at 1st AVR
Valve size at 1st AVR
Underlying disease
Underlying hemodynamic mechanism – AR versus AS versus both
Perfusion time
Concomitant prosthetic MVR
Initial Hospital length of stay
Era effect - < 1990 versus 1990s and up to 2003
Geography – State 
Institution
4.	Causes of death – Big piles only
5.	Early deaths (PCCC) versus late deaths (NDI) – compare profiles
6.	Interactions
7.	Confounders – collinearity
8.	Kaplan Meir – Ross versus Bio versus Mechanical AVR
9.	Competing risk – will think thru this some more
10.	Redo stuff – we may not have time for this before abstract deadline
 
```{r examine-dates}
isolate_month <- function(v) {
  v                                             %>%
    gsub("/.*", "", .)                          %>%
    gsub("-.*", "", .)                          %>%
    return
}
isolate_day <- function(v) {
  v                                             %>%
    sub("^.*?/", "", .)                         %>%
    sub("/.*$?", "", .)                         %>%
    sub("^.*?-", "", .)                         %>%
    sub("-.*$?", "", .)                         %>%
    return
}
isolate_year <- function(v) {
  v                                             %>%
    sub("^.*/", "", .)                          %>%
    sub("^.*-", "", .)                          %>%
    return
}
examine_dates <- function(v) {
  two_slashes <- grepl("/", sub("/", "", v))
  two_dashes  <- grepl("-", sub("-", "", v))
  short_year  <- nchar(isolate_year(v))==2
  special_cases <- 
    ! (two_slashes | two_dashes) |
    short_year                   |
    is.na(v)
  if (sum(special_cases)>0) {
    cat("\n\n")
    cat("Special cases")
    v[special_cases]                            %>%
      table(useNA="ifany")                      %>%
      addmargins                                %>%
      print
  }
  cat("\n\n")
  cat("Month")
  v[!special_cases]                             %>%
    isolate_month                               %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
  cat("\n\n")
  cat("Day")
  v[!special_cases]                             %>%
    isolate_day                                 %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
  cat("\n\n")
  cat("Year")
  v[!special_cases]                             %>%
    isolate_year                                %>%
    as.numeric                                  %>%
    table(useNA="always")                       %>%
    addmargins                                  %>%
    print
}
dlist1 <- c(
  "birthdate",
  "first_admit_date",
  "date_of_valve_replacement",
  "discharge_date",
  "date_of_valve_replacement_1",
  "discharge_date_1",
  "date_of_valve_replacement_2",
  "discharge_date_2",
  "unos_init_date",
  "unos_end_date",
  "unos_tx_date",
  "unos_re_tx_date",
  "pccc_tx_date",
  "ndi_date_of_death")
for (d in dlist1) {
  cat("\n\n*** ")
  cat(d)
  examine_dates(mv1[, d]) 
}
{
  cat("\n\n")
  cat("date_of_death")
  examine_dates(mv3$date_of_death)
}
```

```{r fix_dates}
# change any year that is listed as 1902 to missing (999)
fix_list1 <- "date_of_valve_replacement"
for (d in fix_list1) {
  v <- mv1[, d]
  double_delimiter <-  grepl("[-/].*[-/]", v)
  single_delimiter <-  grepl("[-/]", v) & ! double_delimiter
  absent_delimiter <- !grepl("[-/]", v)
  year_only <- absent_delimiter & as.numeric(v) %in% 1950:2020
  short_year  <- nchar(isolate_year(v))==2
  missing_code <- absent_delimiter & !year_only
  double_00 <- grepl("00/00", v)
  single_00 <- grepl("/00/",  v) & !double_00 
  cat("\n\nMissing code\n")
  v[missing_code] %>% unique %>% sort %>% print
  cat("\n\nYear only\n")
  v[year_only]        %>% unique %>% sort %>% print
  cat("\n\nSingle delimiter\n")
  v[single_delimiter] %>% unique %>% sort %>% print
  cat("\n\nDouble 00\n")
  v[double_00]        %>% unique %>% sort %>% print
  cat("\n\nSingle 00\n")
  v[single_00]        %>% unique %>% sort %>% print
  cat("\n\n")

  w <- v
  yr <- isolate_year()
  w[short_year] <- sub("[/-]..")
  w[year_only] <- paste0("06/30/", v[year_only])
  w[single_delimiter] <- sub("/", "/15/", v[single_delimiter])
  w[single_00] <- sub("/00/", "/15/", v[single_00])
  w[double_00] <- sub("00/00", "06/30", v[double_00])
  print(data.frame(changes=paste(v, "changes to", w)[v!=w]))
}